---
title: "w5codealong"
author: "Teague Tran"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms
ggplot(data = chinstraps, aes(x = body_mass_g)) + 
  geom_histogram(bins = 10) + 
  facet_wrap(~sex)
#QQ plot
ggplot(data = chinstraps, aes(sample = body_mass_g)) + 
      geom_qq() + 
      facet_wrap(~sex)

# get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )

```


Get vectors of body_mass_g for female and male chinstrap penguins separately

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, a couple of different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest


cohen.d(chinstraps_f, chinstraps_m)
```

Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins?

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
```


## one-way ANOVA

### 1. Table of counts by species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)
penguin_counts
```

Create subset of only female penguins

```{r}
penguin_f <- penguins %>%  
  filter(sex == "female")

ggplot(data = penguin_f, aes(x= flipper_length_mm)) +
  geom_histogram(aes(fill = species)) +
  facet_wrap(~species, scales = "free")

ggplot(data = penguin_f, aes(sample= flipper_length_mm)) +
  geom_qq(aes(color = species)) +
  facet_wrap(~species, scales = "free")
```

### summary statistics

find mean of flipper length,
sd
sample size
variance 'var'
standarded error
```{r}
penguin_f_summary<- penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean (flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
    var_flipper_length = var(flipper_length_mm)
  )

leveneTest(flipper_length_mm ~ species, data = penguin_f)
```

### beeswarm plot with mean +- 1sd

```{r}
ggplot() + 
  geom_beeswarm(data = penguin_f, aes(x=species, y= flipper_length_mm,
                                      color = "gray70",
                                      size = 3,
                                      alpha = .5)) +
  geom_point(data = penguin_f_summary, aes(x = species, y = mean_flipper_length,
                                        size = 4)) +
  geom_errorbar(data = penguin_f_summary, aes(x = species,
                                              ymax=mean_flipper_length + sd_flipper_length,
                                              ymin=mean_flipper_length - sd_flipper_length),
                width = 0.1) + 
  theme_light()
```
### one-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov)

TukeyHSD(penguin_aov)
```

Type III anova with car package:

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```
























